---
title: "Global Classements Analysis (all AFTT)"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Global Classements Analysis (all AFTT)}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

## Analysis of current percentage of each classement 

```{r}
library(PingMeUp)
data("players_m", package = "PingMeUp")

library(ggridges)
library(ggplot2)
```

```{r}
# Active players
Actifs_AFTT <- players_m[players_m[, "position_bis"] != "Inactive", ]

# Actives by classement
tab_classement_Actifs_AFTT <- table(Actifs_AFTT[, "classement"])
pct_classement_AFTT <- tab_classement_Actifs_AFTT / nrow(Actifs_AFTT)

df_class_AFTT <- data.frame(
  classement = names(tab_classement_Actifs_AFTT),
  pct_classement_AFTT = round(as.numeric(pct_classement_AFTT),digits=4)
)
noAB0<-c("NC","E6","E4","E2","E0","D6","D4","D2","D0","C6","C4","C2","C0","B6","B4","B2")
df_class_AFTT[df_class_AFTT$classement %in% noAB0,]

# Actives by classement letter
tab_lettre_Actifs_AFTT <- table(Actifs_AFTT[, "classement_lettre"])
pct_letter_AFTT <- tab_lettre_Actifs_AFTT / nrow(Actifs_AFTT)
pct_label_AFTT <- paste0(round(100 * pct_letter_AFTT), "%")

df_labels_AFTT <- data.frame(
  classement_lettre = names(tab_lettre_Actifs_AFTT),
  pct_letter_AFTT = as.numeric(pct_letter_AFTT),
  pct_label_AFTT = pct_label_AFTT
)
df_labels_AFTT
```

Plot of the frequency of each classement for all AFTT players (actives or all)

```{r, fig.cap="graph.pct.classements", fig.align='center', fig.height=5, fig.width=6, out.width='100%'}

graph.pct.classements(players_m)
graph.pct.classements(players_m,actifs_only = FALSE)

```

## Analysis of points per classement

```{r, fig.cap="Points_class"}
players_AFTT_noA <- players_m[players_m$classement_lettre != "A", ]
players_AFTT_noA$classement <- factor(players_AFTT_noA$classement)
Points_class_qt <- aggregate(points ~ classement, data = players_AFTT_noA, FUN = quantile)
Points_class_qt <- do.call(data.frame, Points_class_qt)

Points_class_mean <- aggregate(points ~ classement, data = players_AFTT_noA, FUN = mean)
Points_class <- merge(Points_class_mean, Points_class_qt, by = "classement")
names(Points_class) <- c("classement",
                         "mean_pts",
                         "min_pts",
                         "qt25_pts",
                         "median_pts",
                         "qt75_pts",
                         "max_pts")

Points_class


classement_cols <- c(B0 = "darkred",B2 = "darkred",B4 = "darkred",
  B6 = "darkred",C0 = "darkorange",C2 = "darkorange",
  C4 = "darkorange",C6 = "darkorange",D0 = "goldenrod",
  D2 = "goldenrod",D4 = "goldenrod",D6 = "goldenrod",
  E0 = "olivedrab4",E2 = "olivedrab4",E4 = "olivedrab4",
  E6 = "olivedrab4",NC = "dodgerblue3")

ggplot(players_AFTT_noA,
       aes(x = points,y = classement,
         fill = classement, color = classement)) +
  geom_density_ridges(scale = 2, rel_min_height = 0.01,
                      color = NA) +
  geom_vline(data = Points_class,
    aes(xintercept = mean_pts, color = classement),
    linewidth = 0.4, alpha = 0.5) +
  scale_fill_manual(values = classement_cols) +
  scale_color_manual(values = classement_cols) +
  scale_x_continuous(breaks = Points_class$mean_pts,
                     labels = round(Points_class$mean_pts, 0)) +
  theme_minimal(base_size = 14) +
  xlab("points (1er mois précédent)") +
  theme(axis.text.x = element_text(angle = 90,vjust = 0.5,
                                   hjust = 1,
                                   color = classement_cols[Points_class$classement],face = "bold"),
        legend.position = "none")
```



## Estimate of new classement and analysis of change

To estimate the new classement for a series of players, use the `players.new.classement()` function. It is based on current points of each players and the total number of active players retrieved by `count.actives()`. The function adds 2 columns in the data frame: the new classement (`classement_new`) and the number of classements upward or downward (`classement_diff`). After computing, two tables are displayed, showing
- a transition table with frequencies of players in each pair of old to new classement. Most players (about 50%) don't change of classement hence the diagonal is strong. Also there are more upward changes than downward changes, meaning training efforts generally pays-off! However you will notice that from D4 and upper, there is usually more people going one classement down than up.
- a difference table, where we see most players don't change of classement and only few gain 3 or more classements in a season. The average is around 0.3 classement, which is a rate to which every club or province could compare as a way to measure performance.

Applied to all players (default), the computation gives:
```{r, class.output="small-text"}
count.actives()
players_m_new <- players.new.classement()
summary(players_m_new$classement_diff)
```

```{r, fig.align='center', fig.height=6, fig.width=6, out.width='100%'}
attr(players_m_new, which="diff_table")
plot(attr(players_m_new, which="diff_table")[-length(attr(players_m_new, which="diff_table"))],
     main="Montées et descentes de classement",
     xlab ="Nombre de classements en plus ou moins",
     ylab ="Fréquence (nombre de joueurs)")
```

