---
title: "Global Classements Analysis (all AFTT)"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Global Classements Analysis (all AFTT)}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---
*To be cont'd*

## Analysis of current percentage of each classement 

```{r}
library(PingMeUp)
data("players_m", package = "PingMeUp")
```

```{r}
# Active players
Actifs_AFTT <- players_m[players_m[, "position_bis"] != "Inactive", ]

# Actives by classement
tab_classement_Actifs_AFTT <- table(Actifs_AFTT[, "classement"])
pct_classement_AFTT <- tab_classement_Actifs_AFTT / nrow(Actifs_AFTT)

df_class_AFTT <- data.frame(
  classement = names(tab_classement_Actifs_AFTT),
  pct_classement_AFTT = round(as.numeric(pct_classement_AFTT),digits=4)
)
noAB0<-c("NC","E6","E4","E2","E0","D6","D4","D2","D0","C6","C4","C2","C0","B6","B4","B2")
df_class_AFTT[df_class_AFTT$classement %in% noAB0,]

# Actives by classement letter
tab_lettre_Actifs_AFTT <- table(Actifs_AFTT[, "classement_lettre"])
pct_letter_AFTT <- tab_lettre_Actifs_AFTT / nrow(Actifs_AFTT)
pct_label_AFTT <- paste0(round(100 * pct_letter_AFTT), "%")

df_labels_AFTT <- data.frame(
  classement_lettre = names(tab_lettre_Actifs_AFTT),
  pct_letter_AFTT = as.numeric(pct_letter_AFTT),
  pct_label_AFTT = pct_label_AFTT
)
df_labels_AFTT
```

Plot of the frequency of each classement for all AFTT players (actives or all)

```{r, fig.align='center', fig.height=5, fig.width=6, out.width='100%'}

graph.pct.classements(players_m)
graph.pct.classements(players_m,actifs_only = FALSE)

```

## Estimate of new classement and analysis of change

To estimate the new classement for a series of players, use the `players.new.classement()` function. It is based on current points of each players and the total number of active players `count.actives()`. The function adds 2 columns in the data frame: the new classement (`classement_new`) and the number of classements upward or downward (`classement_diff`). After computing, two tables are displayed, showing
- a transition table with frequencies of players in each pair of old to new classement. Most players (about 50%) don't change of classement hence the diagonal is strong. Also there are more upward changes than downward changes, meaning training efforts generally pays-off!
- a difference table, where we see most players don't change of classement and only few gain 3 or more classements in a season. The average is around 0.3 classement, which is a rate to which every club or province could compare as a way to measure performance.

Applied to all players (default), the computation gives:
```{r, class.output="small-text"}
count.actives()
players_m_new <- players.new.classement()
summary(players_m_new$classement_diff)
```

```{r, fig.align='center', fig.height=6, fig.width=6, out.width='100%'}
attr(players_m_new, which="diff_table")
plot(attr(players_m_new, which="diff_table")[-length(attr(players_m_new, which="diff_table"))],
     main="Montées et descentes de classements",
     xlab ="Nombre de classements en plus ou moins",
     ylab ="Fréquence (nombre de joueurs)")
```

