---
title: "Club and Players Registries"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Club and Players Registries}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

## Overview

The ping package creates two main registries:

-   **Club registry**: a snapshot of all AFTT clubs and their metadata.
-   **Players registry (Messieurs)**: a snapshot of all players with their points and rankings

Each registry can be updated once per month (one recording per month maximum otherwise overwritten) knowing players' points are updated monthly (first days of the month) in the AFTT. AFTT clubs are likely updated yearly.

For both registries a **latest snapshot**, and an **archive** containing all past snapshots are stored.

------------------------------------------------------------------------

## How the registries are built

The registries are generated by running two scripts `build.club.registry.R` and `build.players.registry.R` (available within the `data-raw` folder available on github repository)

`build.club.registry.R` creates a new club registry snapshot from AFTT annuaire `https://data.aftt.be/annuaire/membres.php`. The process involves

- fetching club indices, names, and provinces (one web page)
- scraping postal addresses and team counts from every club's page.
- Merging into a consolidated club registry

`build.club.registry.R` calls the function: `get.club.list.R`

`build.players.registry.R` retrieves the Messieurs player list for every club in current `club_registry`. Iterate through provinces then clubs. Enriches the data with classement letters/numbers separation and within-province ranking.

`build.players.registry.R` calls the functions: `get.club.player.list.m`

------------------------------------------------------------------------

## What the outputs look like

### Club registry

Two datasets are created:

-   `club_registry`\
    The **latest** snapshot (a data frame).

-   `club_registry_archive`\
    A **named list** of past snapshots.\
    Each element is named `"YYYY_MM"` and contains a data frame with the same structure as the latest snapshot.
    
See `club_registry.R`

### Players registry (Messieurs)

Similarly:

-   `players_m`\
    The **latest** snapshot (a data frame).

-   `players_m_archive`\
    A **named list** of past snapshots, also named `"YYYY_MM"`.

------------------------------------------------------------------------

## Loading the registries

All registry datasets are stored in the 'data' folder, included in the installed package and made available to users.

Users can load the latest snapshots with:

```{r}
data("club_registry", package = "PingMeUp")
nrow(club_registry)

#For a specific province
head(club_registry[club_registry$province=="N",])
```


or for players:

```{r}
data("players_m", package = "PingMeUp")
names(players_m)
players_m[order(players_m$position_bis)[1:5],]
```

The archived versions are obtained with

```{r eval=FALSE}
data("club_registry_archive", package = "PingMeUp")
data("players_m_archive", package = "PingMeUp")
```

or for specific snapshots once the dataset is loaded:

```{r eval=FALSE}
club_registry_archive[["2026_04"]]
players_m_archive[["2026_04"]]
```
