---
title: "Players Summaries"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Players Summaries}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---


A number of aggregate outputs are pre-computed with `players.summary.R`

## Total and active players

```{r include=FALSE}
data("players_m", package = "PingMeUp")
```


```{r}
PingMeUp::count.players()
PingMeUp::count.actives()
```
## Active players per province or classement

Actives per province:
```{r}
PingMeUp::table.actives.prov()
```

Actives per classement (detailled or letter)
```{r}
PingMeUp::table.actives.class()
PingMeUp::table.actives.class.l()
#Where you see 75% of players are D,E or NC
round(100*prop.table(PingMeUp::table.actives.class.l()),digits=2)
```

## Wrapping up and subsets:

A wrapper is made to compute those aggregates.
```{r}
PingMeUp::players.summary()
```
 All can be applied to subsets, for example a club:
 
```{r}
PingMeUp::players.summary(players_m[players_m$club=="N051",])
```

a set of provinces, e.g. French speaking:

```{r}
FRprov<-c("BBW","H","L","Lx","N")
PingMeUp::players.summary(players_m[players_m$prov %in% FRprov,])
```

## Number of players and clubs

```{r include=FALSE}
data("players_m", package = "PingMeUp")
data("club_registry", package = "PingMeUp")
```

Number of registered players (active and inactive) per club
```{r}
club_n_players <- data.frame(table(players_m$club))
names(club_n_players) <- c("indice","n_players")
club_n_players <- merge(club_n_players, club_registry[,c("indice","label","province")],by="indice")
summary(club_n_players$n_players)
```

```{r}
library(ggplot2)

ggplot(club_n_players, aes(x = province, y = n_players)) +
  geom_boxplot(fill = "lightblue") +
  labs(
    title = "Distribution of Club Size by Province",
    x = "Province",
    y = "Number of Players"
  ) +
  theme_minimal()
```

 

